---
title: "TidyTemplate"
date: 2021-08-11
output: html_output
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

```

```{r packages}

library(tidyverse)
library(tidytuesdayR)
library(here)
library(forcats)
library(ggtext)
library(ggthemes)
library(patchwork)
library(ggrepel)

```


# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2021-08-10")

investment <- tt$investment
chain_investment <- tt$chain_investment
ipd <- tt$ipd

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

## Explore data

```{r wrangle-data}
top_n_categories <- 4 # Number of categories to include in plot

# Remove the total categories
infrastructure_investments <- chain_investment %>%
  filter(meta_cat != "Total basic infrastructure" &
           meta_cat != "Total infrastructure") %>%
  group_by(meta_cat, year) %>% 
  summarize(inv = sum(gross_inv_chain)) %>%
  filter(year >= 1997) %>% # just show latest 20 years
  mutate(pct_change = (inv - lag(inv, 20)) / lag(inv,20))

biggest_pct_increases <- infrastructure_investments %>%
  ungroup() %>%
  slice_max(order_by = pct_change, n = top_n_categories) %>%
  select(meta_cat)

biggest_pct_decreases <- infrastructure_investments %>%
  ungroup() %>%
  slice_min(order_by = pct_change, n = top_n_categories) %>%
  select(meta_cat)

increase_data <- infrastructure_investments %>% 
  right_join(biggest_pct_increases) %>%
  fill(pct_change, .direction = "up") %>%
  arrange(pct_change, meta_cat, year) %>%
  mutate(label = case_when(
    pct_change > 0 ~ str_c("+", as.character(signif(100*pct_change, digits = 2)), "%"),
    TRUE ~ str_c("-", as.character(signif(100*pct_change, digits = 2)), "%")
    )
  )

decrease_data <- infrastructure_investments %>% 
  right_join(biggest_pct_decreases) %>%
  fill(pct_change, .direction = "up") %>%
  arrange(pct_change, meta_cat, year) %>%
  mutate(label = case_when(
    pct_change > 0 ~ str_c("+", as.character(signif(100*pct_change, digits = 2)), "%"),
    TRUE ~ str_c(as.character(signif(100*pct_change, digits = 2)), "%")
    )
  )

```

## Plot data

```{r plot}
p_increase <- ggplot(data = increase_data,
                     mapping = aes(x = year, y = inv)) +
  geom_line(color = "#084f63",
            size = 1.1,
            show.legend = FALSE, ) +
  geom_label_repel(data = increase_data %>%
                     filter(meta_cat == "Digital" & year == 2017),
                   mapping = aes(x = year, y = inv, label = label),
                   min.segment.length = Inf,
                   nudge_x = -6,
                   size = 2.5) +
  facet_wrap(vars(fct_reorder(meta_cat, pct_change, .desc = TRUE)), 
             nrow = 2, ncol = top_n_categories) +
  scale_y_continuous(breaks = c(0, 50000, 100000, 150000, 200000, 250000), 
                     labels = c("0", "50", "100", "150", "200", "250"),
                     limits = c(0, 255000)) +
  theme_tufte() +
  theme(
    text = element_text(family = "Gill Sans"),
    plot.subtitle = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_rect(color = "white", fill = "gray"),
    plot.background = element_rect(color = "#fefff7", fill = "#fefff7"),
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(color = "#f5f5f5"),
    plot.title.position = "plot"
  ) +
  labs(
    title = "US infrastructure spending from 1997 to 2017 (adjusted for inflation)",
    subtitle = "Categories with *greatest increases* in spending"
  ) +
  ylab("Billions (USD)")

p_decrease <- ggplot(data = decrease_data,
                     mapping = aes(x = year, y = inv)) +
  geom_line(color = "#ab051e",
            size = 1.1,
            show.legend = FALSE) +
  geom_label_repel(data = decrease_data %>%
                     filter(meta_cat == "Public safety" & year == 2017),
                   mapping = aes(x = year, y = inv, label = label),
                   min.segment.length = Inf,
                   size = 2.5) +
  facet_wrap(vars(fct_reorder(meta_cat, pct_change, .desc = TRUE)), 
             nrow = 2, ncol = top_n_categories) +
  scale_y_continuous(breaks = c(0, 50000, 100000, 150000, 200000, 250000), 
                     labels = c("0", "50", "100", "150", "200", "250"),
                     limits = c(0, 255000)) +
  theme_tufte() +
  theme(
    text = element_text(family = "Gill Sans"),
    plot.subtitle = element_markdown(),
    axis.title.y = element_markdown(),
    strip.background = element_rect(color = "white", fill = "gray"),
    plot.background = element_rect(color = "#fefff7", fill = "#fefff7"),
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(color = "#f5f5f5"),
    plot.title.position = "plot",
    plot.caption = element_text(color = "gray60")
  ) +
  labs(
    subtitle = "Categories with *greatest decreases* in spending",
    caption = "@alanjern | Data: Bureau of Economic Analysis"
  ) +
  ylab("Billions (USD)")

p_increase / p_decrease

```

# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(here("figures", "infrastructure.png"))
       

```
